Performance comparison of Dask and Apache Spark on HPC systems

### 2021-07-13 latex/paper.tex
edits and comments in methods

\section{Methods}
\subsection{Infrastructure}
For our experiments we used the SlashBin cluster at Concordia University.
Each compute node has 2 16-cores CPUs,
\SI{256}{\giga\byte} \SI{2666}{\mega\hertz} memory in 8-channel,
and \SI{2.88}{\tera\byte} SSDs of mounted storage.
The nodes are interconnected with a dedicated \SI{10}{\giga\bit/\second} Ethernet connection.
Centos 8 with Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64} installed on the compute nodes.

Both Spark and Dask are configured to have 8 worker processes each with 8 threads.
Each worker is allocated \SI{31.5}{\giga\byte} of memory.
A new cluster is spinned up and teared down for each experiment.

@@

\subsection{Dataset}
We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel intensity ranging between 0 and 65,535.
We converted the blocks into the NIfTI format, a popular format in neuroimaging.
We left the NIfTI blocks uncompressed, resulting in a total data size of \SI{648}{\giga\byte}.
To evaluate the effect of block size, we resplit these blocks into 1000, 2500, and 5000 blocks of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and \SI{129.6}{\mega\byte}, respectively.
\href{https://github.com/big-data-lab-team/sam}{sam} library was used to resplit the image.

We also used the dataset provided by the Consortium for Reliability and Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) \cite{zuo2014open}, freely available on \href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{datalad}.
The entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and functional images of 1,397 subjects acquiredin 29 sites.
We used all 3,491 anatomical images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on average).

\subsection{Applications}
\subsubsection{Increment}
We adapted the increment application used in \cite{hayot2019performance}.
This synthetic application reads blocks of the BigBrain from Lustre and simulates computation by sleeping for a specified period.
To simulate intermediate results, we repeat the sleep process for a configurable amount of time.
We prevent data caching of the blocks by incrementing their voxels value by one after each sleep operation.
Finally, we write the resulting NIfTI image back to Lustre.
This application allows us to study the engines when their inputs are processed independently.
The map-only scenario of this application mimics the processing of multiple independent subjects in parallel.

\begin{figure}[!ht]
	\centering
	\includegraphics[height=\columnwidth,
	angle=0]{figures/increment.png}
	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
	\label{fig:graph-increment}
\end{figure}

@@

\subsubsection{Histogram}
As our third application, we calculate the histogram of the BigBrain image. 
The application reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre.
This map-reduce application has a very high read overwrite ratio.
Moreover, this application requires shuffling, albeit of a limited amoutn of data. 
The amount of inter-worker communication is in-between the increment and multi-increment applications.

@@

\subsubsection{Kmeans}
For our fourth application, we apply Kmeans clustering to the voxel intensities of the BigBrain image.
We set the number of clusters to 3, to segment the white and grey matter and the noise.
The application starts by reading the image blocks, combining all voxels in a 1-D array, and choosing initial centroids using the min, max, and intermediate values.
It assigns each voxel to the centroids it is the closest and updates each centroid by computing the average of the voxels associated with it. 
It repeats the assignment and update steps for a configurable amount of time. 
Finally, the voxels of the image blocks are classified and written back to the file system.
Updating the centroids involves substantial data communication between the workers.

@@

\subsubsection{BIDS App example}
Our fifth application is BIDS App example: a neuroimaging pipeline to measure the brain volume from MRIs.
For this application, we use the CoRR dataset.
The application extracts the brain volume of each participant, then computes the average for each group of participants.
Unlike the other applications, BIDS App example is a command-line executed in a Docker image (bids/example on DockerHub).
We converted the Docker image to a Singularity image for use in HPC environments, \MD{Cite paper on reason why this is done.}
using \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}

@@

\subsection{Experiments}
Table~\ref{table:parameters} the four parameters that are varied throughout the experiments.
We varied (1) the number of workers to assess the scalibility of the scheduler for the engine,
(2) the BigBrain block size in Increment, Multi-Increment, and Histogram to measure the effect of the different I/O pattern and parallelization degrees,
(3) the number of iterations to evaluate the effect of number of task,
and (4) the sleep delay to study the effect of task duration.
It should note that increasing the number of iterations for a given sleep delay also increases the total compute time of an application.

To avoid potential external bias such as caching, background process and network load, we ran the applications in randomized order and cleared the page cache in between every experiments.
Each benchmark was run ten times.

For each run, we measure the makespan of the application as well as the cumulative time spent in the different functions for read, processing, and writing data.
The overhead calculation for each CPU thread is the end time of the last processed task minus the total runtime of the tasks ran for this thread.
Summing those results gives the total overhead for the application.

>>>>

\section{Methods}
\subsection{Infrastructure}
We used the ``slashbin" cluster at Concordia University.
The cluster has 8 compute nodes, 4 storage nodes, 1 login node, and 1
Lustre metadata node. Each
compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
@ 2.10GHz, 256~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
file system (no RAID enabled), 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
CentOS~8.1 and Linux kernel
\textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.

\TG{describe the Lustre installation: data nodes and lustre config}
\TG{Use GB or GiB consistently}

Both Spark and Dask were configured to have 8 worker processes each with 8
threads. Each worker was allocated \SI{31.5}{\giga\byte} of memory \TG{,
determined as the amount of memory that ...}. \TG{any other Spark or Dask
param? Location of log files, etc} A new Dask or Spark cluster was started and
teared down for each experiment.

@@

\subsection{Dataset}
We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel
intensities ranging from 0 to 65,535. We converted the blocks into the
NIfTI format, a popular format in neuroimaging. We left the NIfTI blocks
uncompressed, resulting in a total data size of \SI{648}{\giga\byte}. To
evaluate the effect of block size, we resplit these blocks into 1000, 2500,
and 5000 blocks of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and
\SI{129.6}{\mega\byte}, respectively. The
\href{https://github.com/big-data-lab-team/sam}{sam} library was used to
resplit the image.

We also used the dataset provided by the Consortium for Reliability and
Reproducibility
(\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR})
\cite{zuo2014open}, freely available on
\href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion
and functional images of 1,397 subjects acquired in 29 sites. We used all
3,491 anatomical images, representing \SI{39}{\giga\byte} overall
(\SI{11.17}{\mega\byte} per image on average).

\subsection{Applications}

\subsubsection{Increment}

We adapted the increment application used in \cite{hayot2019performance}.
This synthetic application reads blocks of the BigBrain from Lustre and
simulates computation by sleeping for a specified period. To simulate
intermediate results, we repeat the sleep process for a configurable amount
of time. We prevent data caching of the blocks by incrementing their voxels
value by one after each sleep operation. Finally, we write the resulting
NIfTI image back to Lustre. This application allows us to study the engines
when their inputs are processed independently. The map-only pattern
 mimics the processing of multiple independent subjects in
parallel.

\begin{figure}[!ht]
	\centering
	\includegraphics[height=\columnwidth,
	angle=0]{figures/increment.png}
	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.
	\TG{nitpick: I'd prefer using ``read and write" than ``load and dump''}}
	\label{fig:graph-increment}
\end{figure}

@@

\subsubsection{Histogram}
As our third application, we calculate the histogram of the BigBrain image. 
The application reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre.
This map-reduce application has a very high read overwrite ratio.
Moreover, this application requires shuffling, albeit of a limited amount of data. 
The amount of inter-worker communication is in-between the increment and multi-increment applications.

@@

\subsubsection{Kmeans}
For our fourth application, we apply Kmeans clustering to the voxel
intensities of the BigBrain image. We set the number of clusters to 3. The
application starts by reading the image blocks, combining all voxels in a
1-D array, and choosing initial centroids using the min, max, and
intermediate values. It assigns each voxel to its closest centroid and
updates each centroid with the average of its assigned voxels. It repeats
the assignment and update steps for a configurable number of iterations.
Finally, the voxels of the image blocks are classified and written back to
the file system. Updating the centroids involves substantial data
communication between the workers.

@@

\subsubsection{BIDS App example}
Our fifth application is BIDS App example, a neuroimaging pipeline to
measure the brain volume from MRIs. For this application, we use the CoRR
dataset. The application extracts the brain volume of each participant,
then computes the average for each group of participants. Unlike the other
applications, BIDS App example is a command-line executed in a Docker image
(bids/example on DockerHub). We converted the Docker image to a Singularity
image for use in HPC environments, \MD{Cite paper on reason why this is
done.} using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}

@@

\TG{You should explain how the applications were implemented in Dask and
Spark, that you reused the same code in both  cases, which APIs were used 
and why, etc}

\subsection{Experiments}
Table~\ref{table:parameters} shows the parameters that were varied
throughout the experiments. We varied (1) the number of workers to assess
the scalibility of the engine scheduler, (2) the BigBrain
block size in Increment, Multi-Increment, and Histogram to measure the
effect of different I/O patterns and parallelization degrees, (3) the
number of iterations to evaluate the effect of number of task, and (4) the
sleep delay to study the effect of task duration. It should note that
increasing the number of iterations for a given sleep delay also increases
the total compute time of an application.

To avoid potential external biases due to caching, background processes and
network load, we ran the applications in randomized order and cleared the
page cache in between every experiments. Each benchmark was run ten times.

For each run, we measured the makespan of the application as well as the
cumulative time spent in the different functions for read, processing, and
writing data. The overhead calculation for each CPU thread is the end time
of the last processed task minus the total runtime of the tasks ran for
this thread \TG{unclear definition. Don't refer to CPU threads as a given task may be executed on more than 1. Maybe refer to engine executors?}. Summing those results gives the total overhead for the
application.
